A thesis draft

### 2018-11-29 res/sections/04-Credits.tex

\noindent A big thank for my friends, whose power of cheering me up and make
me smile saved myself different times from the risk of failing this journey.
Thanks because you where always here when I needed you, and you always found the
right words to say, or the right thing to do. Big part of my success is thanks
to them, who have pushed me in places I would not be able to reach alone. In
particular I would like to thank Giovanni, Giorgio, Michele, Gianluca and
Umberto for the good times we end up spending every weekend. On top of that, I
would like to say also thanks to Marco, Federico, Matteo, Mirko, Emanuele,
Stefano, Sara, Lisa and Eduard that closely witnessed with me the good and the
bad moments of the last two years of this journey. The time spent together with
you guys will not be easily forgotten.

>>>>

\noindent A big thank for my friends, whose power of cheer me up and make me
smile saved myself different times from the risk of failing this journey. Thanks
because you where always here when I needed you, and you always found the right
words to say, or the right thing to do. Big part of my success is thanks to
them, who have pushed me in places I would not be able to reach alone. In
particular I would like to thank Giovanni, Stefano, Giorgio, Michele, Gianluca
and Umberto for the good times we end up spending every weekend. On top of that,
I would like to say also thanks to Marco, Federico, Matteo, Mirko, Emanuele,
Stefano (yes, again, you deserve it!), Sara, Lisa and Eduard that closely
witnessed with me the good and the bad moments of the last two years of this
journey. The time spent together with you guys will not be easily forgotten.

### 2018-11-29 res/sections/05-Abstract.tex

\vspace*{\fill}
\par
\begingroup
\leftskip4em
\rightskip\leftskip
\section*{\centering Abstract}
The importance of Internet is growing every year. Mobile and wired traffic are
continuously surging, asking for better performances. On this day, ISPs rely on
optimized, ad-hoc and often in-hardware nework solutions. These network
functions are costly and require time to be deployed and mantained. In this
context, upcoming virtualization technolgies like hypervisor-based and
container-based are providing the possibility to create virtualized network
functions, that have the ability to be up and running in a matter of seconds.
VNFs are then managed and orchestrated by a specific application, the MANO that
handles their lifecycle.

In this thesis we are going to propose an Infrastructure Management and
Orchestrator plane Proof-of-Concept, based on the container virtualization
technology and on the 3GPP Release 15 specifications. In the current PoC, Docker
and the container orchestrator Kubernetes play a fundamntal role.
\par
\endgroup
\vspace*{\fill}
%\bigskip

>>>>

\vspace*{\fill}
\par
\begingroup
\leftskip4em
\rightskip\leftskip
\section*{\centering Abstract}
Network resource management and service optimization are two basic
functionalities widely adopted by network operators. However, provisioning these
two key capability is becoming increasingly difficult considering the high
number of services delivered through today networks. The usual approach of
optimization through middleboxes deployed in the end-to-end path between
user/service is proving incapable in keeping up with the current pace of change.
On the other side, advances in virtualization and softwarization techniques
could provide the much needed flexibility, serving as enablers for an
end-to-end, softwarized network management plane. In this context, taking
inspiration from the recent proposal of the 3GPP Release 15, we propose a
two-layer service orchestration architecture exploiting container technology. In
specifics, we provide a proof of concept implementation of an infrastructure
Management and Orchestration (MANO) plane, replying on the Kubernetes container
orchestration framework.
\par
\endgroup
\vspace*{\fill}
%\bigskip

### 2018-11-29 res/sections/07-Introduction.tex

This thesis is organized as follows:
\begin{itemize}
 \item Chapter~\ref{chap:background} describes background information about 5G
   and the technologies used to build an ETSI compliant architecture;
 \item Chapter~\ref{chap:rel_wk} gives an insight of correlated works on this
   field;
 \item Chapter~\ref{chap:prjan} explains the analysis performed on the
   requirements provided and the design choices performed;
 \item Chapter~\ref{chap:archimpl} provides an in-depth view of the high-level
   architectural implementation;
 \item Chapter~\ref{chap:vnf_ns_impl} gives a brief explanation of the VNF and NS
   implementation;
 \item Chapter~\ref{chap:test} describes the tests done on the platform;
 \item Chapter~\ref{chap:conclusions} wraps up the overall project experience.
\end{itemize}

>>>>

This thesis is organized as follows:
\begin{itemize}
 \item Chapter~\ref{chap:background} describes background information about 5G
   and the technologies used to build an ETSI compliant architecture;
 \item Chapter~\ref{chap:rel_wk} gives an insight of correlated works on this
   field;
 \item Chapter~\ref{chap:prjan} explains the analysis performed on the
   requirements provided and the overall project roadmap;
 \item Chapter~\ref{chap:archimpl} provides an in-depth view of the high-level
   architectural implementation;
 \item Chapter~\ref{chap:vnf_ns_impl} gives a brief explanation of the SFC
   implementation;
 \item Chapter~\ref{chap:test} describes the tests done on the platform;
 \item Chapter~\ref{chap:fw} illustrates possible future works;
 \item Chapter~\ref{chap:conclusions} wraps up the overall project experience.
\end{itemize}

### 2018-11-29 res/sections/08-Background.tex

With the constraints placed on the requirements formulated by the NGMN alliance,
5G envisage a multi-layered architecture, based on three main
layers~\cite{alliance20155g}:
\begin{itemize}
\item \textbf{infrastructure resource layer}: physical resources that are 
exposed via a virtualized interface, and that can be monitored using specific 
APIs
\item \textbf{business enablement layer}: where a library of functions and
  deployment is contained, and its configuration is accessible via APIs
\item \textbf{business application layer}: layer that contains specific
  applications and services of the operator
\end{itemize}

@@

This thesis started as a part of the VIBES project~\cite{vibesesa}, whose main
requirement is to enhance the performance of end-to-end IP based that involves
satellite connections. To reach this goal, the project specifications suggest to
exploit the 5G incoming technology and use the NFV-MANO architecture to perform
first packets elaboration and performance improvement and finally TCP/IP
satellite chunk optimization with the Performance Enhancing Proxy (PEP). The
VIBES project proposed five technical requirements:
\begin{enumerate}
  \item Analysis of the applicability of current and new Internet protocols in
  the proposed VNF-PEP architecture
  \item Implementation of a VNF-PEP prototype
  \item Building of a PoC test platform
  \item VNF-PEP validation and performance tests on 5G use-cases
  \item Demonstration test-bed management
\end{enumerate}

>>>>

With the constraints placed on the requirements formulated by the NGMN alliance,
5G envisage a multi-layered architecture, based on three main
layers~\cite{alliance20155g}:
\begin{itemize}
\item \textbf{infrastructure resource layer}: physical resources that are 
exposed via a virtualized interface, and that can be monitored using specific 
APIs;
\item \textbf{business enablement layer}: where a library of functions and
  deployment is contained, and its configuration is accessible via APIs;
\item \textbf{business application layer}: layer that contains specific
  applications and services of the operator.
\end{itemize}

@@

This thesis started as a part of the VIBES project~\cite{vibesesa}, whose main
requirement is to enhance the performance of end-to-end IP based that involves
satellite connections. To reach this goal, the project specifications suggest to
exploit the 5G incoming technology and use the NFV-MANO architecture to perform
first packets elaboration and performance improvement and finally TCP/IP
satellite chunk optimization with the Performance Enhancing Proxy (PEP). The
VIBES project proposed five technical requirements:
\begin{enumerate}
  \item Analysis of the applicability of current and new Internet protocols in
  the proposed VNF-PEP architecture;
  \item Implementation of a VNF-PEP prototype;
  \item Building of a PoC test platform;
  \item VNF-PEP validation and performance tests on 5G use-cases;
  \item Demonstration test-bed management.
\end{enumerate}
